---
title: "8 Environments"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Environments

```{r include=FALSE}
deck <- read.csv("deck.csv")
```

```{r include=FALSE}
library(pryr)
```
> You can see R’s environment system with the `parenvs` function in the `pryr` package. `parenvs(all = TRUE)` will return a list of the environments that your R session is using.

```{r}
parenvs(all = TRUE)
```

## Working with Environments

__Refer to an environment__

```{r}
as.environment("package:stats")
```
There are some functions for access to `global`, `base` and `empty` environments:

```{r}
globalenv()
baseenv()
emptyenv()
```

__Parent environment__

```{r}
parent.env(globalenv())
```

> Notice that the empty environment is the only R environment without a parent:

```{r}
# parent.env(emptyenv())
# Error in parent.env(emptyenv()) : the empty environment has no parent
```

>You can view the objects saved in an environment with `ls` or `ls.str`. `ls` will return just the object names, but `ls.str` will display a little about each object’s structure:

```{r}
ls(emptyenv())
```

```{r}
ls(globalenv())
```

```{r}
ls.str(globalenv())
```

>You can use R’s $ syntax to access an object in a specific environment. For example, you can access deck from the global environment:

```{r}
head(globalenv()$deck)
```

__`assing` function__

```{r}
assign("number", 100, envir = globalenv()) #  works similar to <-
globalenv()$number
```

### The Active Environment

```{r}
environment()
```

## Scoping Rules

>1. R looks for objects in the current active environment.
2. When you work at the command line, the active environment is the global environment. Hence, R looks up objects that you call at the command line in the global environment.
3. When R does not find an object in an environment, R looks in the environment’s parent environment, then the parent of the parent, and so on, until R finds the object or reaches the empty environment.


## Evaluation

__Runtime environment__

```{r}
show_env <- function(){
  list(
    ran.in = environment(), 
    parent = parent.env(environment()), 
    objects = ls.str(environment())
  )
}

show_env() # R creates a new environment each time you run a function.
```
__Origin environment__

>R will connect a function’s runtime environment to the environment that the function was first created in. Let’s call this environment the *origin environment*. You can look up a function’s origin environment by running environment on the function:

```{r}
environment(show_env)
```

__Objects inside runtime environment__

>Any objects created by the function are stored in a safe, out-of-the-way runtime environment.

```{r}
show_env <- function(){
  text <- "sample text"
  
  list(
    ran.in = environment(), 
    parent = parent.env(environment()), 
    objects = ls.str(environment())
  )
}

show_env()
```

>If a function has arguments, R will copy over each argument to the *runtime environment*.

```{r}
show_env <- function(text){
  list(
    ran.in = environment(), 
    parent = parent.env(environment()), 
    objects = ls.str(environment())
  )
}

show_env("sample text")
```

__Fixing `deal` and `shuffle` functions__

>The following code will save a prisitine copy of deck and then remove the top card:

```{r}
DECK <- deck

deck <- deck[-1, ]
```


```{r}
deal <- function() {
  deck[1, ]
}

environment(deal)
```

>`deal` returns the top card of deck but does not remove the card from the deck. As a result, deal always returns the same card:

```{r}
deal()

deal()
```

```{r}
deal <- function() {
  card <- deck[1, ]
  
  # deck <- deck[-1, ] only create a copy of deck in the function
  # runtime environment
  
  assign("deck", value = deck[-1, ], envir = globalenv())
  card
}

deal()

deal()

deal()
```
>`shuffle` doesn’t shuffle the deck object; it returns a shuffled copy of the deck object:

```{r}
shuffle <- function(cards) { 
  random <- sample(1:52, size = 52)
  cards[random, ]
}
```
```{r}
head(deck, 3)

head(shuffle(deck), 3)

head(deck, 3)
```

Fixing it!

```{r}
shuffle <- function(){
  random <- sample(1:52, size = 52)
  assign("deck", DECK[random, ], envir = globalenv())
}
```

```{r}
shuffle()
```

```{r}
deal()

deal()
```

## Closures

_What happen if `deck` or/and `DECK` are modified or erase from the global environment?_

```{r}
# rm(deck)

# shuffle()
# object 'deck' not foundError in shuffle() : could not find function "shuffle"
```


>It would be better if we could store deck in a safe, out-of-the-way place, like one of those safe, out-of-the-way environments that R creates to run functions in. In fact, storing deck in a runtime environment is not such a bad idea.


```{r}
setup <- function(deck) {
  DECK <- deck

  DEAL <- function() {
    card <- deck[1, ]
    assign("deck", deck[-1, ], envir = globalenv())
    card
  }

  SHUFFLE <- function(){
    random <- sample(1:52, size = 52)
    assign("deck", DECK[random, ], envir = globalenv())
 }

  # return functions as a list
  list(deal = DEAL, shuffle = SHUFFLE)
}
```

> Save each of the elements of the list to a dedicated object in the global environment:

```{r}
cards <- setup(deck)

deal <- cards$deal
shuffle <- cards$shuffle
```

Now, global environment is not longer their _original environment_:

```{r}
environment(deal)

environment(shuffle)
```

>This arrangement is called a closure. setup’s runtime environment “encloses” the deal and shuffle functions. Both deal and shuffle can work closely with the objects contained in the enclosing environment, but almost nothing else can. The enclosing environment is not on the search path for any other R function or environment.

But these functions still modify `deck` and `DECK` in the global environment, another change is needed to solve this:

```{r}
setup <- function(deck) {
  DECK <- deck

  DEAL <- function() {
    card <- deck[1, ]
    assign("deck", deck[-1, ], envir = parent.env(environment()))
    card
  }

  SHUFFLE <- function(){
    random <- sample(1:52, size = 52)
    assign("deck", DECK[random, ], envir = parent.env(environment()))
 }

  list(deal = DEAL, shuffle = SHUFFLE)
}

cards <- setup(deck)
deal <- cards$deal
shuffle <- cards$shuffle
```

Lets see what happen this time if `deck` is erased:
```{r}
rm(deck)

shuffle()

deal()
```
```{r}

```

>We finally have a self-contained card game!
