---
title: "5 R Objects"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Atomic Vectors

- one dimension
- same data type
  - integer
  - double
  - complex
  - character
  - logical
  - raw

__A vector of one integer__
```{r}
number <- 1
is.vector(number)
```
__A vector formed by values of a die__
```{r}
die <- 1:6
```

__A vector of the powers of 10__
```{r}
numbers <- c(1, 10, 100, 1000)
numbers
```

__The length of the above vector__
```{r}
length(numbers)
```

__A vector of one element of type character__
```{r}
str <- "hello"
str
```

### Double
```{r}
x <- 1
typeof(x)
```

```{r}
y <- 0.2
typeof(y)
```

### Integer
>An integer should end with an 'L' in order not to be stored as double
```{r}
week_days <- 7L
typeof(week_days)
```

### Character
```{r}
typeof(str)
```

### Logical
```{r}
is_greater <- 100 > 10
typeof(is_greater)
```

### Null values
```{r}
none <- NULL
typeof(none)
```

## Attributes
__Common Attributes__

- names
- dimensions 
- classes

### Names 
__Set Names__
```{r} 
names(die) <- c("one", "two", "three", "four", "five", "six")
```
__Get Names__
```{r}
# by calling names
names(die)

# by calling attributes
attributes(die)

# by looking at the vector
die
```

__Remove Names__
```{r}
names(die) <- NULL
names(die)
```

### Dim
>Transform an atomic vector into an n-dimensional array.
```{r}
dim(die) <- c(2, 3)
die
```

## Matrices
```{r}
m <- matrix(die, nrow = 2)
m
```
As dim, `matrix` fill the matrix column by column by default. Specifying the argument `byrow = TRUE` is possible to fill the matrix row by row.

```{r}
m <- matrix(die, nrow = 2, byrow = TRUE)
m
```

## Arrays

```{r}
a <- array(die, dim = c(2, 3))
a
```

## Class
```{r}
dim(die) <- c(2, 3)
# type
typeof(die)

# class
class(die)
```

If a object does not have a `class` attribute class will return a value base on the object atomic type.
```{r}
class(number)
class(str)
```

### Factors
R’s way of storing categorical information.

```{r}
color <- factor(c("red", "blue", "green"))
typeof(color)

attributes(color)
```

What is storing R?
```{r}
color
unclass(color)
```

_Factor looks like character but behave like integers._

Treating as character
```{r}
as.character(color)
```

## Coercion
_If a character is present everything will be converted to a character._
```{r}
v <- c(1, 2, 10, "twelve")
v
```

_If a vector only contains logicals and numbers everything will be converted to a numbers._
```{r}
v <- c(TRUE, FALSE, 1L, 0L)
v
```
###  Explicit Conversion
```{r}
as.character(1)

as.numeric("10")

as.logical(0)
```
## Lists
- One-dimensional data set.
- Group R objects like vector or other lists.

```{r}
l <- list(10, "10", c(1:10), FALSE, list("a", "b", "c"))
l

# two-system notation
l[[5]][2]
```
## Data Frames
> Data frames are the two-dimensional version of a list.

```{r}
df <- data.frame(
  names = c("one", "two", "three", "four", "five", "six"),  
  values = c(1:6)
)

df
```

__Data Frame type__
```{r}
typeof(df)
```

__Data Frame class__
```{r}
class(df)
```

__Structure display__
```{r}
str(df)
```

> Data frames store one of the most common forms of data used in data science, tabular data.
## Loading Data
