---
title: "11 Loops"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## expand.grid

>The expand.grid function in R provides a quick way to write out every 
combination of the elements in _n_ vectors.

```{r}
die <- c(1:6)

rolls <- expand.grid(one = die, two = die)
```
```{r}
rolls
```

Determine the value of each roll:

```{r}
rolls$value <- rolls$one + rolls$two

# output of six elements
rolls[sample(c(1:36), 6), ]
```

__Probability of each combination (weighted dice)__

>The probability that n independent, random events all occur is equal to the 
product of the probabilities that each random event occurs.

```{r}
prob <- c("1" = 1/8, "2" = 1/8, "3" = 1/8, "4" = 1/8, "5" = 1/8, "6" = 3/8)
prob
```

Subset by `rolls$one`:
```{r}
prob[rolls$one]
```

Adding the probabilities to the data frame:

```{r}
rolls$p_one <- prob[rolls$one]
rolls$p_two <- prob[rolls$two]

#  probability of each combination
rolls$prob <- rolls$p_one * rolls$p_two

# output of six elements
rolls[sample(c(1:36), 6), ]
```

__Expected value__

```{r}
sum(rolls$value * rolls$prob)
```
> The expected value of rolling two loaded dice is 8.25.

__Expected value of the slot machine prize__

```{r}
wheel <- c("DD", "7", "BBB", "BB", "B", "C", "0")
```


>Add the argument `stringsAsFactors = FALSE` to your `expand.grid` call; 
otherwise, `expand.grid` will save the combinations as factors, an unfortunate 
choice that will disrupt the score function.

```{r}
combinations <- expand.grid(
  first = wheel, 
  second = wheel, 
  third = wheel, 
  stringsAsFactors = FALSE
)

combinations[sample(c(1:nrow(combinations)), 6), ]
```
> Probabilities of each symbol

```{r}
prob <- c(
  "DD" = 0.03,
  "7" = 0.03, 
  "BBB" = 0.06, 
  "BB" = 0.1,
  "B" = 0.25,
  "C" = 0.01,
  "0" = 0.52
)

prob
```

Let's calculate the probability of each combination:

```{r}
combinations$p_first <- prob[combinations$first]
combinations$p_second <- prob[combinations$second]
combinations$p_third <- prob[combinations$third]

#  probability of each combination
combinations$prob <- combinations$p_first * combinations$p_second * combinations$p_third

# output of six elements
combinations[sample(c(1:nrow(combinations)), 6), ]
```

>The sum of the probabilities is one, which suggests that our math is correct:

```{r}
sum(combinations$prob)
```
The remaining step is calculate the expected value; to do this the prize of each
combinations must be determine first with the `score` function.

```{r}
score <- function (symbols) {
  # identify case
  
  # it's possible to test three of the same type with:
  # length(unique(symbols)) == 1
  same <- symbols[1] == symbols[2] && symbols[2] == symbols[3]
  
  bars <- symbols %in% c("B", "BB", "BBB")
  
  # get prize
  if (same) {
    payouts <- c("DD" = 100, "7" = 80, "BBB" = 40, "BB" = 25, 
      "B" = 10, "C" = 10, "0" = 0)
    prize <- unname(payouts[symbols[1]])
  } else if (all(bars)) {
    prize <- 5
  } else {
    cherries <- sum(symbols == "C")
    prize <- c(0, 2, 5)[cherries + 1]
  }
  
  # adjust for diamonds
  diamonds <- sum(symbols == "DD")
  prize * 2 ^ diamonds
}
```

## for Loops

>A for loop repeats a chunk of code many times, once for each element in a 
__set__ of input. 

```{r}
for (v in c(1:6)) {
  print(v)
}
```

Adding a column for prize:

```{r}
combinations$prize <- NA
```

Let's calculate the prize of each row:

```{r}
for (i in 1:nrow(combinations)) {
  symbols <- c(combinations[i, 1], combinations[i, 2], combinations[i, 3])
  combinations$prize[i] <- score(symbols)
}
```


```{r}
sum(combinations$prize * combinations$prob)
```

>This value is lower than expected, because an important feature of the slot 
machine was ignore: a diamond is wild. You can treat a DD as any other symbol 
if it increases your prize, with one exception. You cannot make a DD a C unless 
you already have another C in your symbols.

A version of `score` that handles wild diamonds:

```{r}
score <- function(symbols) {
  
  diamonds <- sum(symbols == "DD")
  cherries <- sum(symbols == "C")
  
  # identify case
  # since diamonds are wild, only nondiamonds 
  # matter for three of a kind and all bars
  slots <- symbols[symbols != "DD"]
  same <- length(unique(slots)) == 1
  bars <- slots %in% c("B", "BB", "BBB")

  # assign prize
  if (diamonds == 3) {
    prize <- 100
  } else if (same) {
    payouts <- c("7" = 80, "BBB" = 40, "BB" = 25,
      "B" = 10, "C" = 10, "0" = 0)
    prize <- unname(payouts[slots[1]])
  } else if (all(bars)) {
    prize <- 5
  } else if (cherries > 0) {
    # diamonds count as cherries
    # so long as there is one real cherry
    prize <- c(0, 2, 5)[cherries + diamonds + 1]
  } else {
    prize <- 0
  }
  
  # double for each diamond
  prize * 2^diamonds
}
```


__Recompute the expected value:__

```{r}
for (i in 1:nrow(combinations)) {
  symbols <- c(combinations[i, 1], combinations[i, 2], combinations[i, 3])
  combinations$prize[i] <- score(symbols)
}
```


```{r}
sum(combinations$prize * combinations$prob)
```

## while Loops

>A `while` loop reruns a chunk while a certain condition remains `TRUE`.

```{r}
i <- 1

while (i <= 10) {
  print(i)
  i <- i + 1
}
```

## repeat Loops

>`repeat` loops are even more basic than `while` loops. They will repeat a chunk
of code until you tell them to stop (by hitting Escape) or until they encounter
the command `break`, which will stop the loop.

```{r}
i <- 1

repeat {
  print(i)
  
  i <- i + 1
  if (i > 10) {
    break
  }
}
```

