---
title: "6 R Notation"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Selecting Values
 
  > You can select values within an R object with R’s notation system.
  
```
object[<subset of rows>, <subset of columns>]
```

 __Ways of write and index in R__
 
 - Positive integers
- Negative integers
- Zero
- Blank spaces
- Logical values
- Names


### Positive Integers
```{r include=FALSE}
deck <- read.csv("../05-R-objects/deck.csv")
```

```{r}
head(deck)

# a single value
deck[1, 1]

# the first row
deck[1, c(1:3)]

# the first column
deck[1:52, 1]

# the first row twice (repeating the index)
deck[c(1,1), 1:3]
```

>R’s notation system is not limited to data frames. You can use the same syntax to select values in any R object, as long as you supply one index for each dimension of the object.


#### Drop

Selecting more than two columns return a __data frame__:
```{r}
new_deck <- deck[1, 1:2]
new_deck
class(new_deck)
```

Selecting a single column return a __vector__:
```{r}
first_column <- deck[1:6, 3]
first_column
class(first_column)
```

By setting `drop = FALSE` is possible to select one column but return a __data frame__:
```{r}
first_column <- deck[1:6, 3, drop = FALSE]
first_column
class(first_column)
```


### Negative Integers

Negative indexes allow for exclusion:
```{r}
# the last row
deck[-(1:51), c(1:3)]
```

> R will return an error if you try to pair a negative integer with a positive integer in the same index:

```{r}
# deck[-(1:51), c(-1, -2, 3)]
# Error in x[j] : only 0's may be mixed with negative subscripts

# the last column of the last row
deck[-(1:51), c(-1, -2)]
```


### Zero

>R will return nothing from a dimension when you use zero as an index.

```{r}
deck[0,0]
```


### Blank Spaces

>You can use a blank space to tell R to extract every value in a dimension.

```{r}
# the first row
deck[1, ]

# the first column
deck[, 1]

# the first six rows
deck[1:6, ]
```


### Logical Values
```{r}
# first two columns of the second row
deck[2, c(TRUE, TRUE, FALSE)]
```


### Names
```{r}
# the first row
deck[1, c(names(deck))]

# first two columns of the second row
deck[2, c("face", "suit")]
```

## Deal and Shuffle the Deck
```{r}
shuffle <- function(deck) {
  ramdom <- sample(1:52, size = 52)
  deck[ramdom, ]
}

deal <- function(deck) {
  shuffle(deck)[1, ]
}
```

```{r}
deal(deck)
deal(deck)
deal(deck)
```
## Dollar Signs and Double Brackets

> Two types of object in R obey an optional second system of notation. You can extract values from data frames and lists with the $ syntax. 

```{r}
values <- deck$value
typeof(values)
```

### $ notation and list

>If you subset a list in the usual way, R will return a new list that has the elements you requested.

```{r}
l <- list(numbers = 1:10, letters = c("a", "b", "c"))
```

__If the list is subset with the usual notation the result will be a new list:__
```{r}
l[1]
typeof(l[1])
```

```{r}
# sum(l[1])
#  Error in sum(l[1]) : invalid 'type' (list) of argument
```

__With the $ notation the values will be selected as they are:__
```{r}
l$numbers
typeof(l$numbers)
```

```{r}
sum(l$numbers)
```

__Using double brackets will do the same as $ notation__
```{r}
l[[1]]
typeof(l[[1]])
```

Double brackets can be combined with the other indexing approaches:
```{r}
l[["numbers"]]
```

